Considering scale in multilayer animal social networks

### 2020-06-15 paper/manuscript.Rmd
reflow, tidy
include mvmt

We generated proximity-based social networks using the R package `spatsoc`
[@Robitaille_2019] in `r version$version.string` [@R_Core_Team_2019]. We assumed
individuals were associating if simultaneous GPS fixes were within the spatial
distance threshold of one another. Typically for ungulates and other gregarious
mammals, the ‘chain rule’ is used for group assignment [@Croft_2008]. For GPS
data, the chain rule is applied by assigning a group identifier to the union of
buffered GPS fixes at each time step. As such, individuals in a group are within
the spatial distance threshold of at least one other though not necessarily all
other individuals [@Robitaille_2019]. Group assignment based on the chain rule
has previously been applied to caribou using a temporal threshold of 5 minutes
and a spatial distance threshold of 50 m [@Peignier_2019; @Lesmerises_2018]. In
all networks, individual caribou were represented as nodes and associations between
individuals in a given network layer were represented as intralayer
edges. <!-- (Figure \@ref(fig:ml)) .-->

@@

Multilayer networks represent a novel framework for explicitly testing the
influence of habitat selection on sociality across scales. Animal social systems
can be parsed by spatial contexts, for example as defined by habitat types as in
Figure \@ref(fig:lcres). Given that habitat data are available across a wide
range of spatial scales, we evaluated the influence of habitat spatial
resolution on multilayer network metrics. The results indicate varied influence
of increasing spatial resolution depending on the habitat layer. As spatial
resolution increased, edge overlap decreased in forage while remaining stable in
open and forest habitats (Figure \@ref(fig:socres) D). Forage is a rare and
patchy habitat on Fogo Island, as is indicated by a decrease in the proportion
of relocations in this habitat type with increasing spatial resolution (Figure
S1)<!--TODO-->. Similar to increasing spatial distance threshold, the influence
of increasing spatial resolution on graph strength was notably different between
habitat layers. Forest layers showed little change in graph strength with
increasing spatial resolution and variation between individuals. Forage and open
layers had greater individual variation in graph strength, while forage was the
sole layer to show large responses in graph strength to increasing spatial
resolution (Figure \@ref(fig:socres) E). As such, multilayer network measures
related to habitat are explicitly dependent on spatial scale. Edge overlap can
indicate if some associations between individuals are dependent on the
habitat-context [@Battiston_2014]. If the spatial resolution of the habitat data
are insufficient to capture rare habitats, context-specific relationships
between individuals may not be captured by multilayer networks.

@@

<!-- Multilayer network extensions:
movement

Given the existing framework for constructing multilayer networks from
behavioural layers [e.g. @Smith_Aguilar_2019], monolayer networks that explicitly
quantify movement parameters could be integrated. For example, edges could
represent measures of dyadic or collective movement based on dynamic interaction
metrics [@Long_2014]. Integrating movement layers with habitat-specific social
association layers could further illuminate the role of movement as a driver
of the social and spatial environments an individual experiences [@Webber_2018].
In our case study, we generated multilayer networks based on habitat-specific
monolayers. Coordinated movement of dyads, or groups, could vary based on
habitat, and the approach we outline could be used to explore mechanisms linking
the social and spatial environments. For example, social processes may be an
emergent property of the landscape because animals aggregate at resources, such
as waterholes [@Chamaill_Jammes_2008]. Movement to, from, or within habitat
patches or territories can also contribute to the formation of the social
environment [@Spiegel_2016]. For example, in sleepy lizards (*Tiliqua rugosa*)
males interacted primarily at home range edges, presumably driving territorial
behaviour [@Spiegel_2018]. Incorporating movement ecology within a multilayer
network framework is a logical next step for a burgeoning field. In particular,
when association networks are habitat-specific, we suggest the integration of
movement and space use in multilayer networks provides novel insights into the
effects of habitat configuration on the formation of the social environment
[@He_2019; @Webber_2018].

GPS caveat HMM

Multilayer networks can be built with different types of interactions, a differ
In contrast to multilayer networks formed by combining layers defined
by different types of interactions, multilayer networks with GPS data are limited to 
proximity based measures. An alternative which has been explored is using 
hidden markov models behavioural states, within which individuals are associating. [@Jones_2020]

t bon

-->

<!-- put this in -->
Importantly, the key to incorporating movement in multilayer networks across
scales is the technological advancements available to overcome previous
limitations [@Hughey_2018]. Biologging and wildlife tracking technology has
advanced to the point where relocations can be programmed to occur as frequently
as each second [@B_rger_2016], a practice which has been applied to questions of
collective movement [e.g @Strandburg_Peshkin_2015]. Meanwhile, remote sensing
technology, including unmanned aerial vehicles and satellites, can be used to
track the fine-scale movements of groups over large distances and time scales
[@Torney_2018]. Using technology to infer behaviour has therefore become
standard operating procedure for many species in many systems. Our ability to
collect and process remotely sensed data and infer social processes based on
animal movement provides an opportunity to use multilayer social network
analysis to answer complex questions about animal socioecology.

>>>>

We generated proximity-based social networks using the R package `spatsoc`
[@Robitaille_2019] in `r version$version.string` [@R_Core_Team_2019]. We assumed
individuals were associating if simultaneous GPS fixes were within the spatial
distance threshold of one another. Typically for ungulates and other gregarious
mammals, the ‘chain rule’ is used for group assignment [@Croft_2008]. For GPS
data, the chain rule is applied by assigning a group identifier to the union of
buffered GPS fixes at each time step. As such, individuals in a group are within
the spatial distance threshold of at least one other though not necessarily all
other individuals [@Robitaille_2019]. Group assignment based on the chain rule
has previously been applied to caribou using a temporal threshold of 5 minutes
and a spatial distance threshold of 50 m [@Peignier_2019; @Lesmerises_2018]. In
all networks, individual caribou were represented as nodes and associations between
individuals in a given network layer were represented as intralayer
edges.

@@

Multilayer networks represent a novel framework for explicitly testing the
influence of habitat selection on sociality across scales. Animal social systems
can be parsed by spatial contexts, for example as defined by habitat types as in
Figure \@ref(fig:lcres). Given that habitat data are available across a wide
range of spatial scales, we evaluated the influence of habitat spatial
resolution on multilayer network metrics. The results indicate varied influence
of increasing spatial resolution depending on the habitat layer. As spatial
resolution increased, edge overlap decreased in forage while remaining stable in
open and forest habitats (Figure \@ref(fig:socres) D). Forage is a rare and
patchy habitat on Fogo Island, as is indicated by a decrease in the proportion
of relocations in this habitat type with increasing spatial resolution (Figure
S1)<!--TODO-->. Similar to increasing spatial distance threshold, the influence
of increasing spatial resolution on graph strength was notably different between
habitat layers. Forest layers showed little change in graph strength with
increasing spatial resolution and variation between individuals. Forage and open
layers had greater individual variation in graph strength, while forage was the
sole layer to show large responses in graph strength to increasing spatial
resolution (Figure \@ref(fig:socres) E). As such, multilayer network measures
related to habitat are explicitly dependent on spatial scale. Edge overlap can
indicate if some associations between individuals are dependent on the
habitat-context [@Battiston_2014]. If the spatial resolution of the habitat data
are insufficient to capture rare habitats, context-specific relationships
between individuals may not be captured by multilayer networks.
<!-- --- -->

Another extension of multilayer networks using GPS location data is integrating
movement. Given the existing framework for constructing multilayer networks from
behavioural layers [e.g. @Smith_Aguilar_2019], monolayer networks that explicitly
quantify movement parameters could be integrated. For example, edges could
represent measures of dyadic or collective movement based on dynamic interaction
metrics [@Long_2014]. Integrating movement layers with habitat-specific social
association layers could further illuminate the role of movement as a driver
of the social and spatial environments an individual experiences [@Webber_2018].
We generated multilayer networks based on habitat-specific layers of 
social association. Coordinated movement of dyads, or groups, could vary based on
habitat, and the approach we outline could be used to explore mechanisms linking
the social and spatial environments. For example, social processes may be an
emergent property of the landscape because animals aggregate at resources, such
as waterholes [@Chamaill_Jammes_2008]. Movement to, from, or within habitat
patches or territories can also contribute to the formation of the social
environment [@Spiegel_2016]. For example, in sleepy lizards (*Tiliqua rugosa*)
males interacted primarily at home range edges, presumably driving territorial
behaviour [@Spiegel_2018]. Incorporating movement ecology within a multilayer
network framework is a logical next step for a burgeoning field. In particular,
when association networks are habitat-specific, we suggest the integration of
movement and space use in multilayer networks provides novel insights into the
effects of habitat configuration on the formation of the social environment
[@He_2019; @Webber_2018].

Importantly, the key to incorporating movement in multilayer networks across
scales is the technological advancements available to overcome previous
limitations [@Hughey_2018]. Biologging and wildlife tracking technology has
advanced to the point where relocations can be programmed to occur as frequently
as each second [@B_rger_2016], a practice which has been applied to questions of
collective movement [e.g @Strandburg_Peshkin_2015]. Meanwhile, remote sensing
technology, including unmanned aerial vehicles and satellites, can be used to
track the fine-scale movements of groups over large distances and time scales
[@Torney_2018]. Using technology to infer behaviour has therefore become
standard operating procedure for many species in many systems. Our ability to
collect and process remotely sensed data and infer social processes based on
animal movement provides an opportunity to use multilayer social network
analysis to answer complex questions about animal socioecology.

@@

<!--
Multilayer networks can be built with different types of interactions

In contrast to multilayer networks formed by combining layers defined
by different types of interactions, multilayer networks with GPS data are limited to 
proximity based measures. An alternative which has been explored is using 
hidden markov models behavioural states, within which individuals are associating. [@Jones_2020]

netTS
-->
